Experiment 3 of 4

### 2016-07-28 experiments/e3/writeup/experiment3.Rmd
tidying

---
title: "Experiment 3 of 4"
output:
  html_document:
    toc: yes
---  
# Experiment 3

```{r, startClean}
#Start clean.
rm(list=ls())
```

```{r,docOpts}
#Knitr doc options.
library(knitr)
#knit_hooks$set(crop=hook_pdfcrop)
#opts_chunk$set(cache=TRUE, autodep=TRUE)
#options(tidy=TRUE)
#options(replace.assign=TRUE, width=40)
```

```{r, loadLibraries}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

@@

```{r,sourceFunctions}
source("functions/CIfunctions.R")
```

@@

```{r,readdata}
dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)

@@

# # print to file a table with information about the design
# design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Selection, Quantity, Order, 
#                                            Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
# design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
# row.names(design_info) <- NULL
# capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")

# # put dat in better column order
# dat <- subset(dat, 
#               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Selection, 
#                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
#                          c_Trl, s_Trl, c_Itm, c_Vag, Selection.sum, f_Cnd, c_Ord, c_Qty,
#                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
#                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
#                          Instruction, nchar_instr
#               ))

@@

save(dd, file="dd.Rda")

```

@@

```{r}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r}
load("dd.Rda")
suppressPackageStartupMessages(require(lme4))
suppressPackageStartupMessages(require(lmerTest))
suppressPackageStartupMessages(require(languageR))
```

```{r}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r}
summary(v5)
```

```{r}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

@@

```{r}
par(las=2, mar=c(14,6,1,1))
y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5)[-c(1,2,3),]
y=cbind(y,q)
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
grid()
```

@@

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

>>>>

---
title: "Experiment 3 of 4"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  
# Experiment 3

```{r docOpts}
library(knitr)
```

```{r loadLibraries}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(require(languageR))
```

@@

```{r sourceFunctions}
source("functions/CIfunctions.R")
```

@@

```{r readdata}
dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)

@@

@@

# save dd as dd.Rda
# This data set non-impossible trials and renamed/recomputed variables
save(dd, file="dd.Rda")
```

@@

```{r boxplot subjects and items}
load("dd.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r load_dd}
load("dd.Rda")
```

```{r do_lmer}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r summary_lmer}
summary(v5)
```

```{r R2}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r plotLMER}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

@@

```{r coeff}
par(las=2, mar=c(14,6,1,1))
y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5)[-c(1,2,3),]
y=cbind(y,q)
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
grid()
```

@@

```{r baayenplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism
